Anony*mice*d shareable data: Using *mice* to create and analyze multiply imputed synthetic data sets

### 2021-10-26 manuscript/Manuscript.Rmd
Start revising paper.

---
title: "Anony*mice*d shareable data: Using *mice* to create and analyze multiply imputed synthetic data sets"
author:
  - name: Thom Volker
    affil: 1,*, \dagger
    orcid: 0000-0002-2408-7820
  - name: Gerko Vink
    affil: 1, \dagger
    orcid: 0000-0001-9767-1924
affiliation:
  - num: 1
    address: |
      Utrecht University - 
      Department of Methodology and Statistics. 
      Padualaan 14. 3584CH Utrecht, the Netherlands
# firstnote to eighthnote
firstnote: |
  These authors contributed equally to this work.
correspondence: |
  t.b.volker@uu.nl.
journal: psych
type: article
status: submit
bibliography: synth.bib
appendix: appendix.tex
abstract: |
  Synthetic data sets can greatly improve the dissemination of data 
  and further analysis of private data. Generating and analyzing synthetic data 
  sets is straightforward, yet adopting a synthetic data analysis pipeline is 
  often ignored by applied researchers. We outline a simple procedure for 
  generating and alayzing synthetic data sets with `mice` in `R`. We demonstrate 
  in simulation that the analysis results obtained on synthetic data 
  obtained with `mice` yields unbiased and valid inferences, and leads to synthetic 
  records that cannot be distinguished from the true data records. The ease of use 
  when synthesizing data with `mice`, together with the validity of inferences 
  obtained through this procedure opens up a wealth of possibilities for data 
  dissemination and further research of initially private data. 
keywords: |
  mice; imputation; synthetic data.
acknowledgement: |
  We are grateful to Mirthe Hendriks and Stijn van den Broek for replicating the 
  simulation study on an independent data set. 
authorcontributions: |
  The authors contributed equally to this work.
conflictsofinterest: |
  The authors declare no conflict of interest. 
sampleavailability: |
  A full simulation archive is available from https://github.com/amices/Federated_imputation/tree/master/manuscript.
output: rticles::mdpi_article
---

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions on data privacy and data confidentiality. Although these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of modern day social science research [@lazer_life_2009].

@@

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic data sets allow for correct statistical inferences, despite the fact that the analyses do not use the "true" value. The analyses over multiple synthetic data sets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

To mitigate additional effort of creating synthetic data sets on behalf of the researcher, software aimed at multiple imputation of missing data can be employed. Especially if researchers used this software at an earlier stage in the research process, or acquired familiarity with it during earlier projects, the additional burden of creating synthetic data sets is relatively small. The R-package `mice` [@mice] implements multiple imputation of missing data in a straightforward and user-friendly manner. However, the functionality of `mice` is not restricted to the imputation of missing data, but allows to impute any value in the data: even observed values. Consequently, `mice` can be utilized for the creation of multiply imputed synthetic data sets.

After creating the multiply imputed synthetic data sets, the goal is to obtain valid statistical inferences in the spirit of @rubin_multiple_1987 and @neyman1934. In the missing data framework, this is done by performing statistical analyses on all imputed data sets, and pooling the results of the analyses according to Rubin's rules [@rubin_multiple_1987, pp.76]. In the synthetic data framework, the same procedure is followed, but with a slight twist: there are no values that remain constant over the synthetic data sets. The procedure of drawing valid inferences from multiple synthetic data sets is therefore slightly different.

@@

The `mice` package [@mice] in `R` [@Rproject] has been developed for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missings. Doing so, `mice` makes use of fully conditional specification [FCS; @vanbuuren_fully_2006], which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, where $k$ denotes the number of columns in the data. Using FCS, a model is constructed for every incomplete variable and the missing values $Y_{j, mis}$ are then imputed with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable-by-variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed data sets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$.

In `mice`, the generation of multiply imputed data sets to solve for unobserved values is straightforward. The following pseudocode details the multiple imputation of the `mice::boys` data set [@fredriks_boys_2000] into the object `imp` with `m = 10` imputated sets and `maxit = 7` iterations for the algorithm to converge, using the default imputations methods for each column data class.

@@

It is straightforward to extended the imputation approach to generate synthetic values. Rather than imputing missing data, the observed values are then replaced by synthetic draws from the posterior predictive distribution. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let for $n$ units denote $Z_i = 1$ if any of the values of unit $i = 1, 2, \dots, n$ are to be replaced by imputations, and $Z_i = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})$. Now, instead of imputing $\textbf{Y}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, $\textbf{Y}_{rep}$ is imputed from the posterior distribution of $P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)$, where $l$ is an indicator for the synthetic data set ($l = 1, 2, \dots, m$). Note that synthetic values that are imputed at an earlier step can be used for imputing variable $j$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$.

For example, overimputing synthetic values for both the observed and missing cells in the `mice::boys` data set into the object `syn`, given the same imputation parameters as before, can be realized by the following code execution.

@@

where the argument `where` requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix) containing logicals $z_{ij}$ that indicate which cells are selected to have their values replaced by draws from the posterior predictive distribution. This approach allows to *overimpute* a subset of the observed data, or - as in the above example - the observed data as a whole, resulting in a data set that partly or completely consists of synthetic data values.

Choosing an adequate imputation model to impute the data is paramount, as a flawed imputation model may drastically impact the validity of inferences. Imputation models should be as flexible as possible to capture most of the patterns in the data, and to model possibly unanticipated data characteristics [@murray_multiple_2018; @rubin_18years_1996]. Parametric methods, albeit easy to implement in practice, may be too restrictive to capture generally complex patterns in the data, especially in the case of nonlinear relations and interactions between multiple variables. Classification and regression trees [CART; @breiman_cart_1984] allow to model more complex patterns in the data, and have therefore been suggested as an appropriate imputation method [@reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014]. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we will focus our illustrations and evaluations of `mice` to method `mice.impute.cart()`, realized by:

@@

We demonstrate the suitability of using `mice` for synthesization using a simulation study on the `mice::boys` data set. This data set consists of the values of $`r nrow(mice::boys)`$ Dutch boys on the following $`r ncol(mice::boys)`$ variables:

@@

Unfortunately, this data set does not differ from the vast majority of collected data sets, in the sense that it suffers from missing data. For simplicity, we complete the missing values using the default `mice` imputation model for all predictors except `bmi`, which is passively imputed using its deterministic relation with `wgt` and `hgt`.

@@

To induce sampling variance, 1000 bootstrap samples of the `boys` data have been synthesized with $m = 5$ imputations for every data cell. Synthetic values are generated using the `CART` imputation method for all columns, except for `bmi`. The deterministic relation `bmi` which will be synthesized passively based on the synthetic values for `hgt` and `wgt` to preserve the relation in the synthetic data. Additional parameters that come with the use of `mice.impute.cart()` are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both constrain the flexibility of the imputation model. The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. In the current example, the complexity parameter is specified at `cp = 1e-08` rather than the default value `1e-04`, and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic data set is reduced. Additionally, since the synthesis pattern is monotone, the number of iterations can be set to `maxit = 1`.

To assess the performance of `mice` for synthesizing data, we compare the bootstrapped samples with the synthetic versions of these bootstrapped samples. Specifically, univariate descriptive statistics, the correlation matrix, and two linear regression models as well as one ordered logistic regression model will be considered. Subsequently, the bias in the parameters and the $95\%$ confidence interval coverage of the synthetic data will be examined. Similarly to multiple imputation of missing data, correct inferences from synthetic data requires correct pooling over the multiply imputed data sets.

@@

Similarly to the missing data case, variances, and subsequently confidence intervals, should incorporate the increase in variance that is due to imputation [@reiter_partially_inference_2003; @drechsler_synthetic_2011]. Yet, the increase in variance due to imputation differs according to whether missing values are imputed or observed data is overimputed with missing values. Whereas the variance estimate after imputation of missing data needs to account for the fact that a certain amount of information in the data is missing, variance estimation from synthetic data does not suffer from this issue. The adjusted variance estimate that follows from using multiple synthetic data sets only suffers from the fact that a finite number of $m$ synthetic data sets are used to resemble the observed data. Hence, the according variance estimate for synthetic data as developed by @reiter_partially_inference_2003 yields
$$
T = \bar{u}_m + \frac{b_m}{m},
$$

@@

# Results
```{r simulation, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy = FALSE, echo = FALSE}
plan(multisession) # increase speed through futures

@@

## Univariate estimates
The univariate descriptives for the original data and the synthetic data can be found in Table 2. 
```{r univariate-statistics, echo = FALSE, cache = TRUE, results = FALSE}
library(xtable)
true <- psych::describe(data)

synth <- synthetic_samples %>% 
  map(~.x %>% complete("all") %>% 
        map(~.x %>% describe()) %>% Reduce("+", .) / synthetic_samples[[1]]$m) %>% 
  Reduce("+", .) / length(synthetic_samples)
rownames(true) <- paste("original", rownames(true), sep = " ")
rownames(synth) <- paste("synthetic", rownames(synth), sep = " ")

bind_rows(true, setNames(synth, names(true))) %>% 
           arrange(vars) %>% .[, c(2:5, 8:9, 11:12)] %>% round(3) %>% 
  xtable
```
<!-- % latex table generated in R 4.1.0 by xtable 1.8-4 package -->
<!-- % Mon Oct  4 16:47:59 2021 -->
\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}
We see from Table 1 that the synthetic data estimates closely resemble the true data estimates.

## Bivariate estimates
An often used bivariate statistic is Pearson's correlation coefficient. When evaluating this correlation coefficient on the numeric columns in the `boys` data set, we find that biases are very small. The results are displayed in Table 3. 
```{r bivariate-statistics, echo = FALSE, cache = TRUE, results = FALSE}
true <- data %>% select(-gen, -phb, -reg) %>% cor

@@

cbind(boot %>% select(true, boot.bias, boot.cov),
      synth %>% select(synth.bias, synth.cov)) %>% xtable(digits = 3)
```
First, we evaluate the performance of our synthetic simulation set on a linear model where `hgt` is modeled by a continuous predictor `age` and an ordered categorical predictor `phb`. The results for this simulation can be found in Table 4.

>>>>

---
title: "Anony*mice*d shareable data: Using *mice* to create and analyze multiply imputed synthetic data sets"
author:
  - name: Thom Volker
    affil: 1,*, \dagger
    orcid: 0000-0002-2408-7820
  - name: Gerko Vink
    affil: 1, \dagger
    orcid: 0000-0001-9767-1924
affiliation:
  - num: 1
    address: |
      Utrecht University - 
      Department of Methodology and Statistics. 
      Padualaan 14. 3584CH Utrecht, the Netherlands
# firstnote to eighthnote
firstnote: |
  These authors contributed equally to this work.
correspondence: |
  t.b.volker@uu.nl.
journal: psych
type: article
status: submit
bibliography: synth.bib
appendix: appendix.tex
abstract: |
  Synthetic data sets can greatly improve the dissemination of data 
  and further analysis of private data. Generating and analyzing synthetic data 
  sets is straightforward, yet a synthetic data analysis pipeline is 
  seldom adopted by applied researchers. We outline a simple procedure for 
  generating and analyzing synthetic data sets with `mice` in `R`. We demonstrate 
  using simulations that the analysis results obtained on synthetic data 
  yields unbiased and valid inferences, and leads to synthetic 
  records that cannot be distinguished from the true data records. The ease of use 
  when synthesizing data with `mice`, together with the validity of inferences 
  obtained through this procedure opens up a wealth of possibilities for data 
  dissemination and further research of initially private data. 
keywords: |
  mice; imputation; synthetic data.
acknowledgement: |
  We are grateful to Mirthe Hendriks and Stijn van den Broek for replicating the 
  simulation study on an independent data set. Additionally, we thank Hanne Oberman, [[NOT ANONYMOUS REVIEWER]] and an anonymous reviewer for their stimulating comments.
authorcontributions: |
  The authors contributed equally to this work.
conflictsofinterest: |
  The authors declare no conflict of interest. 
sampleavailability: |
  A full simulation archive is available from https://github.com/amices/Federated_imputation/tree/master/manuscript.
output: rticles::mdpi_article
---

# ADD REVIEWER NAME

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Disseminating the research data bears the potential of increasing the scientific returns for the same data collection effort tremendously. Additionally, the fact that public funds are used for data collection results in increasing demand for access to the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions on data privacy and data confidentiality. Although these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of modern day social science research [@lazer_life_2009].

@@

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily replace any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic data sets allow for correct statistical inferences, despite the fact that the analyses do not use the "true" value. The analyses over multiple synthetic data sets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. The quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses [@reiter2004simultaneous; @grund2021using; @jiang2021balancing]. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, additional complications to actually create these high-quality data sets should be kept at a minimum.

To mitigate unnecessary challenges related to creating synthetic data sets on behalf of the researcher, software aimed at multiple imputation of missing data can be employed. Especially if researchers acquired familiarity with this software during earlier projects, or used it earlier during the research process, the additional burden of creating synthetic data sets is relatively small. The R-package `mice` [@mice] implements multiple imputation of missing data in a straightforward and user-friendly manner. However, the functionality of `mice` is not restricted to the imputation of missing data, but allows to impute any value in the data: even observed values. Consequently, `mice` can be used for the creation of multiply imputed synthetic data sets.

After creating multiply imputed synthetic data sets, the goal is to obtain valid statistical inferences in the spirit of @rubin_multiple_1987 and @neyman1934. In the missing data framework, this is done by performing statistical analyses on all imputed data sets, and pooling the results of the analyses according to Rubin's rules [@rubin_multiple_1987, p. 76]. In the synthetic data framework, the same procedure is followed, but with a slight twist: there are no values that remain constant over the synthetic data sets. The procedure of drawing valid inferences from multiple synthetic data sets is therefore slightly different.

@@

The `mice` package [@mice] in `R` [@Rproject] has been developed for multiple imputation to overcome problems related to nonresponse. In that context, the aim is to replace missing values with plausible values from the posterior predictive distribution of that variable. In `mice`, this is accomplished using fully conditional specification (FCS) [@vanbuuren_fully_2006], which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, where $k$ denotes the number of columns in the data. Using FCS, a model is constructed for every incomplete variable and the missing values $Y_{j, mis}$ are then imputed with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable-by-variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed data sets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$.

In `mice`, the generation of multiply imputed data sets to solve for unobserved values is straightforward. The following pseudocode details the multiple imputation of the `mice::boys` data set [@fredriks_boys_2000] into the object `imp` with `m = 10` imputated sets and `maxit = 7` iterations for the algorithm to converge, using the default imputations methods given each column's data class.

@@

It is straightforward to extend the imputation approach to generate synthetic values. Rather than imputing missing data, the observed values are then replaced by synthetic draws from the posterior predictive distribution. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let, given $n$ observations, $Z_i = 1$ if any of the values of unit $i = 1, 2, \dots, n$, are to be replaced by imputations, and $Z_i = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})$. Now, instead of imputing $\textbf{Y}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, $\textbf{Y}_{rep}$ is imputed from the posterior distribution of $P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)$, where $l$ is an indicator for the synthetic data set ($l = 1, 2, \dots, m$). The synthetic values that are imputed at an earlier step can be used for imputing variable $j$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$. Note that, given a completely observed data set, iterating over the posterior predictive distribution is not required.

If the goal is to create synthetic versions of a data set that suffers from missingness, the imputation procedure becomes more complex. A two-step imputation procedure has been proposed by @reiter2004simultaneous. In the first step, the missing data is imputed according to the principles of multiple imputation for missing data, while in a second step, the synthetic data is imputed conform the corresponding procedure. Intuitively, these steps could be combined using `mice`, conditioning the imputations on both the missingness and the values that are to be synthesized. In practice, replacing observed and missing cells in the `mice::boys` data set by synthetic values into the object `syn`, given the same imputation parameters as before, can be realized by the following code execution.

@@

Here, the argument `where` requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix) containing logicals $z_{ij}$ that indicate which cells are selected to have their values replaced by draws from the posterior predictive distribution. This approach allows to replace a subset of the observed data, or as in the aforementioned example, the observed data as a whole, resulting in a data set that partially or completely consists of synthetic data values. However, because the validity of this approach when the data suffers from missingness has not been investigated yet, we restrict ourselves in the following sections to a completely observed data set.However, because the validity of this approach has not been investigated yet, we restrict ourselves in the followi

Choosing an adequate imputation model to impute the data is paramount, as a flawed imputation model may drastically impact the validity of inferences [@grund2021using; @jiang2021balancing]. Imputation models should be as flexible as possible to capture most of the patterns in the data, and to model possibly unanticipated data characteristics [@murray_multiple_2018; @rubin_18years_1996]. Parametric methods, albeit easy to implement in practice, may be too restrictive to capture generally complex patterns in the data, especially in the case of nonlinear relations and interactions between multiple variables. Classification and regression trees [CART\; @breiman_cart_1984] allow to model more complex patterns in the data, and have therefore been suggested as an appropriate imputation method [@reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014]. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we will focus our illustrations and evaluations of `mice` to method `mice.impute.cart()`, realized by:

@@

We demonstrate the suitability of using `mice` for synthetisation using a simulation study on the `mice::boys` data set. This data set consists of the values of $`r nrow(mice::boys)`$ Dutch boys on the following $`r ncol(mice::boys)`$ variables:

@@

Unfortunately, this data set does not differ from the vast majority of collected data sets, in the sense that it suffers from missing data. For simplicity, we complete the missing values using the default `mice` imputation model for all predictors except `bmi`, which is passively imputed using its deterministic relation with `wgt` and `hgt`. Specifically, the imputed values are used to calculate the exact `bmi` values that correspond with `hgt` and `wgt`.

@@

Usually, one would draw samples from a population that can be synthesized to evaluate the performance of the synthesization methods. As we only have access to a single sample, 1000 bootstrap samples of the `boys` data have been synthesized with $m = 5$ imputations for every data cell to induce an appropriate amount of sampling variance. Synthetic values are generated using the `CART` imputation method for all columns, except for `bmi`. The deterministic relation `bmi` which will be synthesized passively based on the synthetic values for `hgt` and `wgt` to preserve the relation in the synthetic data. Additional parameters that come with the use of `mice.impute.cart()` are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both constrain the flexibility of the imputation model. The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. In the current example, the complexity parameter is specified at `cp = 1e-08` rather than the default value `1e-04`, and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic data set is reduced. Additionally, since the synthesis pattern is monotone, the number of iterations can be set to `maxit = 1` [e.g., @drechsler_synthetic_2011, Ch. 3].

To assess the performance of `mice` for synthesizing data, we compare the bootstrapped samples with the synthetic versions of these bootstrapped samples. Specifically, univariate descriptive statistics, the correlation matrix, and two linear regression models as well as one ordered logistic regression model will be considered. Subsequently, the bias in the parameters and the $95\%$ confidence interval coverage of the synthetic data will be examined. Similar to multiple imputation of missing data, correct inferences from synthetic data require correct pooling over the multiply imputed data sets.

@@

Also similarly to the missing data case, variances, and subsequently confidence intervals, should incorporate the increase in variance that is due to imputation [@reiter_partially_inference_2003; @drechsler_synthetic_2011]. Yet, the increase in variance due to imputation differs according to whether missing values are imputed or observed data is replaced by synthetic values. Whereas the variance estimate after imputation of missing data needs to account for the fact that a certain amount of information in the data is missing, variance estimation from synthetic data does not suffer from this issue. The adjusted variance estimate that follows from using multiple synthetic data sets only suffers from the fact that a finite number of $m$ synthetic data sets are used to resemble the observed data. Hence, the according variance estimate for synthetic data as developed by @reiter_partially_inference_2003 yields
$$
T = \bar{u}_m + \frac{b_m}{m},
$$

@@

# Results

```{r simulation, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy = FALSE, echo = FALSE}
plan(multisession) # increase speed through futures

@@

## Univariate estimates

The univariate descriptives for the original data and the synthetic data can be found in Table 2.

```{r univariate-statistics, echo = FALSE, cache = TRUE, results = FALSE}
library(xtable)
true <- psych::describe(data)

synth <- synthetic_samples %>% 
  map(~.x %>% complete("all") %>% 
        map(~.x %>% describe()) %>% Reduce("+", .) / synthetic_samples[[1]]$m) %>% 
  Reduce("+", .) / length(synthetic_samples)
rownames(true) <- paste("observed", rownames(true), sep = " ")
rownames(synth) <- paste("synthetic", rownames(synth), sep = " ")

bind_rows(true, setNames(synth, names(true))) %>% 
           arrange(vars) %>% .[, c(2:5, 8:9, 11:12)] %>% round(3) %>% 
  xtable
```
<!-- % latex table generated in R 4.1.0 by xtable 1.8-4 package -->
<!-- % Mon Oct  4 16:47:59 2021 -->
\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations. Variable names followed by a $^*$ are categorical.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}
We see from Table 1 that the synthetic data estimates closely resemble the true data estimates. All sample statistics of interest show negligible bias over the 1000 synthetic data sets. Hence, univariately the imputation model proves adequate.

## Bivariate estimates

An often used bivariate statistic is Pearson's correlation coefficient. When evaluating this correlation coefficient on the numeric columns in the `boys` data set, we find that biases are very small. The results are displayed in Table 3.

```{r bivariate-statistics, echo = FALSE, cache = TRUE, results = FALSE}
true <- data %>% select(-gen, -phb, -reg) %>% cor

@@

cbind(boot %>% select(true, boot.bias, boot.cov),
      synth %>% select(synth.bias, synth.cov)) %>% xtable(digits = 3)
```

First, we evaluate the performance of our synthetic simulation set on a linear model where `hgt` is modeled by a continuous predictor `age` and an ordered categorical predictor `phb`. The results for this simulation can be found in Table 4.

\begin{table}[ht]
\caption{Simulation results for a linear regression model with continuous and ordered categorical predictors. The model evaluated is \texttt{hgt $\sim$ age + phb}. Depicted are the true data estimate and the bias from the true data estimate and the coverage rate of the 95\% confidence interval for the bootstrap and synthetic data sets.}
\centering
\begin{tabular}{rrrrrr}
  \hline
  & & \multicolumn{2}{c}{Bootstrap} & \multicolumn{2}{c}{Synthetic}\\
 term & estimate & bias & cov & bias & cov \\ 
  \hline
(Intercept) & 63.087 & -0.001 & 0.970 & 0.405 & 0.958 \\ 
  age & 7.174 & 0.000 & 0.958 & -0.033 & 0.947 \\ 
  phb.L & -12.250 & 0.008 & 0.950 & 0.582 & 0.927 \\ 
  phb.Q & -1.376 & -0.022 & 0.926 & 0.112 & 0.934 \\ 
  phb.C & -3.564 & 0.051 & 0.915 & 0.301 & 0.912 \\ 
  phb\verb|^|4 & -0.431 & 0.016 & 0.930 & 0.106 & 0.940 \\ 
  phb\verb|^|5 & 2.064 & 0.060 & 0.941 & 0.077 & 0.943 \\ 
   \hline
\end{tabular}
\end{table}
